---
title: "Homework - 1"
output: md_document
---

## Question - 1
```{r, include=FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2)


xtabs(~Origin, data=ABIA)
xtabs(~UniqueCarrier, data = ABIA)

#Let us try to minimize delays by understanding delays in travel when different carriers are involved.
# In other words, which carrier is the worst to travel with?

carrierDest = ABIA %>%
  group_by(UniqueCarrier, Dest) %>%
  summarise(mean_dep_delay = mean(DepDelay, na.rm=TRUE), 
            mean_arr_delay = mean(ArrDelay, na.rm=TRUE), count = n())
```
```{r message=FALSE, warning=FALSE}

ggplot(carrierDest) + 
  geom_col(aes(x = factor(Dest), y = mean_dep_delay)) + 
  facet_wrap(~UniqueCarrier)
```

```{r message=FALSE, warning=FALSE}
ggplot(carrierDest) + 
  geom_col(aes(x = factor(Dest), y = mean_arr_delay)) + 
  facet_wrap(~UniqueCarrier)
```

According to the graph above, we noticed that mean delays were off the charts with the carriers OH and OO. So we'll focus on the reason behind the delays. If the reason is weather (which cant be controlled - except NAS), we do not consider the carrier as one of the worst. 


#### OH - Jetsream Intl.
```{r message=FALSE, warning=FALSE}
OH_delays = ABIA %>%
  group_by(UniqueCarrier, Origin) %>%
  filter (UniqueCarrier == "OH") %>%
  summarise(meanCarDel = mean(CarrierDelay, na.rm = TRUE),
            meanWeaDel = mean(WeatherDelay, na.rm = TRUE),
            meanNASDel = mean(NASDelay, na.rm = TRUE),
            meanSecDel = mean(SecurityDelay, na.rm = TRUE),
            meanLADel = mean(LateAircraftDelay, na.rm = TRUE), count = n())
```
```{r}


ggplot(OH_delays) + 
  geom_col(aes(x = Origin, y = meanCarDel)) +
  labs(title = "Delays due to the carrier")

ggplot(OH_delays) + 
  geom_col(aes(x = Origin, y = meanWeaDel)) +
  labs(title = "Delays due to the Weather")

ggplot(OH_delays) + 
  geom_col(aes(x = Origin, y = meanNASDel)) +
  labs(title = "NAS Delays")

ggplot(OH_delays) + 
  geom_col(aes(x = Origin, y = meanSecDel)) +
  labs(title = "Delays due to the Security")

ggplot(OH_delays) + 
  geom_col(aes(x = Origin, y = meanLADel)) +
  labs(title = "Delays due to the Late Arrivals")
```


#### OO - Skywest Airlines

```{r message=FALSE, warning=FALSE}
OO_delays = ABIA %>%
  group_by(UniqueCarrier, Origin) %>%
  filter (UniqueCarrier == "OO") %>%
  summarise(meanCarDel2 = mean(CarrierDelay, na.rm = TRUE),
            meanWeaDel2 = mean(WeatherDelay, na.rm = TRUE),
            meanNASDel2 = mean(NASDelay, na.rm = TRUE),
            meanSecDel2 = mean(SecurityDelay, na.rm = TRUE),
            meanLADel2 = mean(LateAircraftDelay, na.rm = TRUE), count = n())
```
```{r message=FALSE, warning=FALSE}
ggplot(OO_delays) + 
  geom_col(aes(x = Origin, y = meanCarDel2)) + 
  labs(title = "Delays due to the carrier")

ggplot(OO_delays) + 
  geom_col(aes(x = Origin, y = meanWeaDel2)) +
  labs(title = "Delays due to the Weather")

ggplot(OO_delays) + 
  geom_col(aes(x = Origin, y = meanNASDel2)) +
  labs(title = "NAS Delaysr")

ggplot(OO_delays) + 
  geom_col(aes(x = Origin, y = meanSecDel2)) +
  labs(title = "Delays due to the Security")

ggplot(OO_delays) + 
  geom_col(aes(x = Origin, y = meanLADel2)) +
  labs(title = "Delays due to the Late Arrivals")
```


Looks like weather delays were not the primary reason for delays out of the five, for the two carriers.

This helps us realize that the worst Carriers to travel with are 'OH'- "Jetstream Intl" and 'OO' - "Skywest Airlines" as the delays which are not inevitable are the highest in these two airlines.


#### Now we will try to find out which are the best/worst months to travel in.

```{r}
monthly_del = ABIA %>%
  group_by(Month) %>%
  summarize(mean_dep_delay = mean(DepDelay, na.rm=TRUE), 
            mean_arr_delay = mean(ArrDelay, na.rm=TRUE), count = n())

```
```{r}
ggplot(monthly_del) +
  geom_col(aes(x = Month, y = mean_dep_delay)) + 
  labs(title = "Mean Departure Delay by month")

ggplot(monthly_del) +
  geom_col(aes(x = Month, y = mean_arr_delay)) + 
  labs(title = "Mean Arrival Delay by month")

```


According to the graphs above, the best months to travel in are - September, October and November as the delays are the least during this time.


## Question - 2

####    Top 10 songs of all time
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)

top_10 = billboard %>%
  group_by(song, performer) %>%
  filter(week_position == 1) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  head(10) 

top_10
```
```{r message=FALSE, warning=FALSE}
diversity_table = billboard %>%
  group_by(year, song, performer) %>%
  filter(year != 1958, year != 2021) %>%
  summarize(count = n())


yearly_diversity = diversity_table %>%
  group_by(year) %>%
  summarize(count = n())
```
```{r}
ggplot(yearly_diversity) + 
  geom_line(aes(x = year, y = count)) + 
  labs(title = "Musical Diversity", y = "no. of songs in that year")

```

The graph suggests that musical diversity was the highest in the 60s, which declined and dipped in the 90s and 2000s. The diversity is climbing up steeply again.

```{r}
big_hits = billboard %>%
  group_by(song, performer) %>%
  summarize(count = n())

ten_week_hits = big_hits %>%
  filter(count >= 10) %>%
  group_by(Artist = performer) %>%
  summarize(Count = n()) %>%
  filter(Count>=30) %>%
  arrange(desc(Count))
```
```{r}
ggplot(ten_week_hits) +
  geom_col(aes(x = Artist, y = Count)) +
  coord_flip() +
  labs(title = "30 X 10-Week-Hits by Artist")
```



## Question - 3

95th quantile for all females across all Athletics events:
```{r}
olympics_top20 %>%
  group_by(sex) %>%
  filter(sex == "F", sport == "Athletics") %>%
  summarize(quantile(height, 0.95))
```

#### Variation in women's heights by event:
```{r}

fem_sd_height = olympics_top20 %>%
  group_by(event) %>%
  filter(sex == "F") %>%
  summarise(sd_height = sd(height)) %>%
  arrange(desc(sd_height))

fem_sd_height

```

According to this table, the event with the highest variability in female heights is: Rowing Women's Coxed Fours

```{r}
avg_age = olympics_top20 %>%
  group_by(year) %>%
  filter(sport == "Swimming",) %>%
  summarize(avg_age = mean(age)) %>%
  arrange(year)

ggplot(avg_age) +
  geom_line(aes(x = year, y = avg_age)) +
  labs(title = "Change in average age of Swimmers")

```

This graph was true for all swimmers Now let us look at each sex individually.

```{r}
male_avg_age = olympics_top20 %>%
  group_by(year) %>%
  filter(sport == "Swimming", sex == "M") %>%
  summarize(male_avg_age = mean(age)) %>%
  arrange(year)

fem_avg_age = olympics_top20 %>%
  group_by(year) %>%
  filter(sport == "Swimming", sex == "F") %>%
  summarize(fem_avg_age = mean(age)) 

combined_avg_age = merge(male_avg_age, fem_avg_age, by.x = "year")

ggplot(combined_avg_age) + 
  geom_line(aes(x = year, y = male_avg_age), colour = "darkred") +
  geom_line(aes(x = year, y = fem_avg_age), colour = "steelblue") +
  labs(title = "Change in average age of Male and Female Swimmers", x = "Year", y = "Average Age")
```

The average age of male swimmers seems to have fallen and is gradually rising. The average age of female swimmers, however started low and is now gradually rising.



## Question - 4
```{r}
library(tidyverse)
library(rsample)
library(caret)
library(foreach)

trim_350 = sclass %>%
  filter(trim == "350") %>%
  summarize(mileage, price)

trim_350_split = initial_split(trim_350, prop = 0.8)
trim_350_train = training(trim_350_split)
trim_350_test = testing(trim_350_split)

k_grid = c(2, 4, 6, 8, 10, 15, 20, 25, 30, 35, 40, 45,
           50, 60, 70, 80, 90, 100, 125, 150, 175, 200)

trim350_grid = foreach(k = k_grid, .combine = 'rbind') %dopar% {
  knn = knnreg(price ~ mileage, data=trim_350_train, k = k)
  err = modelr::rmse(knn, trim_350_test)
  c(k=k, err = err)
} %>% as.data.frame

ggplot(trim350_grid) +
  geom_point(aes(x=k, y=err)) + 
  labs(x = "K", y = "RMSE", title = "RMSE~K for sclass 350's")
```

#### The statistics for the least RMSE for 350's are:
```{r}
trim350_grid %>% filter(err == min(trim350_grid$err))

```

```{r}
trim_350_test = trim_350_test %>%
  mutate(price_pred = predict(knnreg(price ~ mileage, data=trim_350_train, k = 20), trim_350_test))

ggplot(data = trim_350_test) + 
  geom_point(aes(x = mileage, y = price)) +
  geom_line(aes(x = mileage, y = price_pred), colour = "darkred") + 
  labs(title = "Observed Values and Fitted line for price~mileage - 350's")
```

```{r}
trim_65amg = sclass %>%
  filter(trim == "65 AMG") %>%
  summarize(mileage, price)

trim_65amg_split = initial_split(trim_350, prop = 0.8)
trim_65amg_train = training(trim_65amg_split)
trim_65amg_test = testing(trim_65amg_split)

trim65_grid = foreach(k = k_grid, .combine = 'rbind') %dopar% {
  knn = knnreg(price ~ mileage, data=trim_65amg_train, k = k)
  err = modelr::rmse(knn, trim_65amg_test)
  c(k=k, err = err)
} %>% as.data.frame

ggplot(trim65_grid) +
  geom_point(aes(x=k, y=err)) +
  labs(x = "K", y = "RMSE", title = "RMSE~K for sclass 65 AMG's")
```

#### The statistics for the least RMSE for 65-AMG's are:
```{r}
trim65_grid %>% filter(err == min(trim65_grid$err))
```
```{r}
trim_65amg_test = trim_65amg_test %>%
  mutate(price_pred = predict(knnreg(price ~ mileage, data=trim_350_train, k = 15), trim_65amg_test))

ggplot(data = trim_65amg_test) + 
  geom_point(aes(x = mileage, y = price)) +
  geom_line(aes(x = mileage, y = price_pred), colour = "darkred") + 
  labs(title = "Observed Values and Fitted line for price~mileage - 65 AMG's")
```

#### The 350's leads to a larger optimal value of 'k'.
Optimal K-value for:
350's       = 20
65-AMG's    = 15
